Differentiation By Year

### 2017-08-17 doc/RMD/by-year.Rmd
use xvalDapc for Year analysis

```{r dapc}
setPop(dat11cc) <- ~Region/Year
regyear.dapc <- dapc(dat11cc, n.pca = 20, n.da = 20)
```

@@

par(mfrow = c(2, 1))
loadingplot(regyear.dapc$var.contr, axis = 1)
loadingplot(regyear.dapc$var.contr, axis = 2)
par(mfrow = c(1, 1))

@@

The loading plot shows us the variables that are important for the first axis of
separation, mainly:

----------------
 Locus    Allele
-------- -------
7-2(F)       174
8-3(H)       256
9-2(F)       372
55-4(F)      157
110-4(H)     370
----------------

@@

```{r fig.width = 10, fig.height = 12}
dat11ccloc <- dat11cc[loc = c("7-2(F)", "8-3(H)", "9-2(F)", "55-4(F)", "110-4(H)")]
loc <- map(seploc(dat11ccloc), . %>% 
             tab() %>% 
             as.data.frame() %>% 
             bind_cols(strata(dat11cc)) %>% 
             select(-MCG) %>%
             group_by(Region, Year) %>% 
             mutate(N = n()) %>%
             filter(N > 5) %>%
             summarize_if(is.numeric, mean, na.rm = TRUE) %>%
             ungroup() %>%
             mutate(Year = as.integer(as.character(Year))) %>%
             gather(allele, frequency, -Region, -Year, -N) %>%
             mutate(allele = as.integer(gsub("^.+?\\.([[:alnum:]]+)$", "\\1", allele)))
             # ,
             # .id = "Locus"
           )

@@

```{r NEplot, fig.width = 10, fig.height = 12}
gglNE <- suppressMessages(ggloc(loc, "NE", 2003:2012))
cowplot::plot_grid(plotlist = gglNE, labels = names(gglNE), ncol = 2)
```
```{r MIplot, fig.width = 10, fig.height = 12}
gglMI <- suppressMessages(ggloc(loc, "MI", 2003:2012))
cowplot::plot_grid(plotlist = gglMI, labels = names(gglMI), ncol = 2)
```

>>>>

```{r dapc, cache = TRUE}
setPop(dat11cc) <- ~Region/Year
set.seed(2017-08-18)
regyear.dapc <- xvalDapc(tab(dat11cc), pop(dat11cc), n.pca.max = 20, n.rep = 1000)$DAPC
regyear.dapc
```

@@

quart <- quantile(regyear.dapc$var.contr, 0.95)
par(mfrow = c(2, 1))
ax1 <- loadingplot(regyear.dapc$var.contr, axis = 1, threshold = quart)$var.names %>%
  strsplit("\\.") %>% 
  map_chr(1) %>%
  unique()
ax2 <- loadingplot(regyear.dapc$var.contr, axis = 2, threshold = quart)$var.names %>%
  strsplit("\\.") %>% 
  map_chr(1) %>%
  unique()
par(mfrow = c(1, 1))

@@

The loading plot shows us the variables that are important for the first axis of
separation.

@@

```{r fig.width = 12, fig.height = 14}
dat11ccloc <- dat11cc[loc = c(ax1, ax2)]
loc <- map(seploc(dat11ccloc), . %>% 
             tab() %>% 
             as.data.frame() %>% 
             bind_cols(strata(dat11cc)) %>% 
             select(-MCG) %>%
             group_by(Region, Year) %>% 
             mutate(N = n()) %>%
             filter(N > 5) %>%
             summarize_if(is.numeric, mean, na.rm = TRUE) %>%
             ungroup() %>%
             mutate(Year = as.integer(as.character(Year))) %>%
             gather(allele, frequency, -Region, -Year, -N) %>%
             mutate(allele = as.integer(gsub("^.+?\\.([[:alnum:]]+)$", "\\1", allele)))
             # ,
             # .id = "Locus"
           )

@@

```{r NEplot, fig.width = 12, fig.height = 14}
gglNE <- suppressMessages(ggloc(loc, "NE", 2003:2012))
cowplot::plot_grid(plotlist = gglNE, labels = names(gglNE), ncol = 2)
```
```{r MIplot, fig.width = 10, fig.height = 12}
gglMI <- suppressMessages(ggloc(loc, "MI", 2003:2012))
cowplot::plot_grid(plotlist = gglMI, labels = names(gglMI), ncol = 2)
```
